\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing income inequality with tax data - Switzerland from 1917 to 2010}

### 2014-12-07 paper/subdocuments/results.Rnw
grosses update. 4er und 6er plot. update von 2010 auf 2011

<<setup_data_Brühlhart with and without taxed>>=
# Ohne Nuller
bd <- read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd$Jahr <- as.numeric(substr(bd$Veranlagungsperiode,1,4))
bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] <- bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] - 2
bd<-bd %>% 
  filter(Einheit=="Total")
bd<-subset(bd,Jahr<1995 | Jahr >2002)
bd.long<-reshape(bd,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income (only taxed)","Equivalised net income (only taxed)"),
direction ="long")
# mit Nuller
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.0<-bd.0 %>% 
  filter(Einheit=="Total")
bd.0<-subset(bd.0,Jahr<1993 | Jahr >2002)
bd.0.long<-reshape(bd.0,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income","Equivalised net income"),
direction ="long")
bdlong<-rbind(bd.0.long,bd.long)
names(bdlong)[61]<-"Income"
@

<<equivalencescale, fig.cap='Gini coefficients with and without equivalization',fig.pos='H'>>=
number_ticks <- function(n) {function(limits) pretty(limits, n)}
fig_equiv <- ggplot(bdlong, aes(x=Jahr,y=Gini,shape=Income))+
  scale_y_continuous(limits=c(0.28,0.55))+
  #scale_x_continuous(limits=c(1970,2012),breaks=number_ticks(10)) +
  scale_x_continuous(limits=c(1970,2011),breaks=number_ticks(10)) +
  geom_point(aes(shape=Income),size=3)+
  geom_line(data=bdlong,aes(linetype=Income))+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Using (pseudo) equivalence scales")
fig_equiv
# Gini Range
max(bd %>% mutate(d_steink=abs(gini_steink-gini_steinka),d_reink=abs(gini_reink-gini_reinka)) %>% select(starts_with("d_")),na.rm=TRUE)
# 0.011
@

@@

<<setup_data_Brülhart_2010>>=
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.2010<-bd.0 %>% 
  filter(Jahr==2010,Einheit=="Total")
start <- which(names(bd.2010)=="p1")
end <- which(names(bd.2010)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
names(bd.2010)[1]=c("inc")
bd.2010 <- rbind(c(0,0),bd.2010)
ival<-bd.2010$inc
perval<-bd.2010$percentile
bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
  len=(perval[i]-perval[i-1])*100
  seq(ival[i-1],ival[i],length.out=len)
}))
bd.2010.ind<-bd.2010.ind/1000
@

@@

# Export data to use with Stata
bd2010_2003 <- rbind(data.frame(cbind("bd" = bd.2010.ind,"group"=2010)),data.frame("bd" = bd.2003.ind,"group"=2003))
write.dta(bd2010_2003, file="../../data/stata_data/bd2010_2003.dta")
@

@@

% Variante 2010
<<setup_data_specialcases10>>=
#library(foreign)
#library(ggplot2)
#library(reldist)
normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2010)
normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
start <- which(names(normal)=="p1")
end <- which(names(normal)=="p99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
normal <- data.frame(t(normal[1,start:end]),percentile)
names(normal)[1]=c("inc")

#bruelhart data
pooled <- filter(read.csv("../../data/data_Schweiz.csv",sep=";"), Veranlagungsperiode==2010, Einheit=="Total")
start <- which(names(pooled)=="p1")
p95 <- which(names(pooled)=="p95")
end <- which(names(pooled)=="p99")
pooled <- data.frame(t(pooled[1,c(start:p95,end)]),percentile)
names(pooled)[1]=c("inc")
pooled$inc<-pooled$inc/1000 #adjust scale
normal$case <- "normal case"
pooled$case <- "pooled, normal and special cases"
percentiles <- rbind(normal,pooled)

@@

<<specialcases10, fig.cap='Comparison of normal and special cases 2010',fig.pos='H'>>=

@@

# Export data to use with Stata
ns2010 <- rbind(data.frame(cbind("income" = pooled_i,"group"=0)),data.frame("income" = normal_i,"group"=1))
ns2010$group <- factor(ns2010$group,labels=c("pooled","normal"))
write.dta(ns2010, file="../../data/stata_data/ns2010.dta")

#Gini Range
gini_allcases <- gini(ns2010$income[ns2010$group=="pooled"]); gini_allcases
gini_normal <- gini(ns2010$income[ns2010$group=="normal"]); gini_normal
round(gini_allcases-gini_normal,3)
# 0.02

@@

# Export data to use with Stata
habe_bd_2010_married <- rbind(data.frame(cbind("income" = habe10.married$inc_habe,"group"=0,"weight"=habe10.married$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
habe_bd_2010_married$group <- factor(habe_bd_2010_married$group,labels=c("habe","bd"))
write.dta(habe_bd_2010_married, file="../../data/stata_data/habe_bd_2010_married.dta")

@@

estv_masse_long$measure <- factor(estv_masse_long$variable)
estv_masse_long$Year <- as.double(as.character(estv_masse_long$Year))
estv_masse_long <- estv_masse_long %>% group_by(measure) %>% mutate(log_value=log(value)-log(value)[1])
fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log_value,shape=measure))+
  #scale_y_continuous(limits=c(0,0.55))+
  scale_x_continuous(limits=c(1917,2011),breaks=number_ticks(10)) +
  geom_point(aes(shape=measure),size=3)+
  geom_line(aes(linetype=measure))+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Using different inequality measures")
fig_measures

>>>>

<<setup_data_Brühlhart with and without taxed>>=
# Ohne Nuller
bd <- read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd$Jahr <- as.numeric(substr(bd$Veranlagungsperiode,1,4))
bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] <- bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] - 2
bd<-bd %>% 
  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_reink,gini_reinka)
bd<-subset(bd,Jahr<1995 | Jahr >2002)
# 2011 zufügen
bd.2011 <- read.csv("../../data/NP_2011.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011)
levels(bd.0$Veranlagungsperiode)<- c(levels(bd.0$Veranlagungsperiode),"2011")
bd <- rbind(bd,bd.2011[,names(bd)])
bd.long<-reshape(bd,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income (only taxed)","Equivalised net income (only taxed)"),
direction ="long")
# mit Nuller
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.0<-bd.0 %>% 
  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_reink,gini_reinka)
bd.0<-subset(bd.0,Jahr<1993 | Jahr >2002)
# 2011 zufügen
bd.0.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011)
levels(bd.0$Veranlagungsperiode)<- c(levels(bd.0$Veranlagungsperiode),"2011")
bd.0 <- rbind(bd.0,bd.0.2011[,names(bd.0)])
bd.0.long<-reshape(bd.0,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income","Equivalised net income"),
direction ="long")
bdlong<-rbind(bd.0.long,bd.long)
names(bdlong)[3]<-"Income"
@

<<equivalencescale, fig.cap='Gini coefficients with and without equivalization',fig.pos='H'>>=
number_ticks <- function(n) {function(limits) pretty(limits, n)}
fig_equiv <- ggplot(bdlong, aes(x=Jahr,y=Gini,shape=Income))+
  scale_y_continuous(limits=c(0.28,0.55))+
  #scale_x_continuous(limits=c(1970,2012),breaks=number_ticks(10)) +
  scale_x_continuous(limits=c(1970,2011),breaks=number_ticks(10)) +
  geom_point(aes(shape=Income),size=3)+
  geom_line(data=bdlong,aes(linetype=Income))+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Using (pseudo) equivalence scales")
fig_equiv
# Gini Range
max(bd %>% mutate(d_reink=abs(gini_reink-gini_reinka)) %>% select(starts_with("d_")),na.rm=TRUE)
max(bd.0 %>% mutate(d_reink=abs(gini_reink-gini_reinka)) %>% select(starts_with("d_")),na.rm=TRUE)
# 0.012
@

@@

<<setup_data_Brülhart_2010>>=
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(Einheit=="Total")
#bd.2010<-bd.0 %>% 
#  filter(Jahr==2010,Einheit=="Total")
#start <- which(names(bd.2010)=="p1")
#end <- which(names(bd.2010)=="p99_99")
#percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
#bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
#names(bd.2010)[1]=c("inc")
#bd.2010 <- rbind(c(0,0),bd.2010)
#ival<-bd.2010$inc
#perval<-bd.2010$percentile
#bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
#  len=(perval[i]-perval[i-1])*100
#  seq(ival[i-1],ival[i],length.out=len)
#}))
#bd.2010.ind<-bd.2010.ind/1000

start <- which(names(bd.2011)=="p1")
end <- which(names(bd.2011)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2011<- data.frame(t(bd.2011[1,start:end]),percentile)
names(bd.2011)[1]=c("inc")
bd.2011 <- rbind(c(0,0),bd.2011)
ival<-bd.2011$inc
perval<-bd.2011$percentile
bd.2011.ind<-unlist(sapply(2:length(ival),function(i){
  len=(perval[i]-perval[i-1])*100
  seq(ival[i-1],ival[i],length.out=len)
}))
bd.2011.ind<-bd.2011.ind/1000
@

@@

# Export data to use with Stata
bd2011_2003 <- rbind(data.frame(cbind("bd" = bd.2011.ind,"group"=2011)),data.frame("bd" = bd.2003.ind,"group"=2003))
write.dta(bd2011_2003, file="../../data/stata_data/bd2011_2003.dta")
@

@@

% Variante 2011
<<setup_data_specialcases11>>=
#library(foreign)
#library(ggplot2)
#library(reldist)
normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2011)
normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
start <- which(names(normal)=="p1")
end <- which(names(normal)=="p99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
normal <- data.frame(t(normal[1,start:end]),percentile)
names(normal)[1]=c("inc")

#bruelhart data
#pooled <- filter(read.csv("../../data/data_Schweiz.csv",sep=";"), Veranlagungsperiode==2010, Einheit=="Total")
pooled <- read.csv("../../data/NP_2011.csv",dec=",") %>% filter(Einheit=="Total")
start <- which(names(pooled)=="p1")
p95 <- which(names(pooled)=="p95")
end <- which(names(pooled)=="p99")
pooled <- data.frame(t(pooled[1,c(start:p95,end)]),percentile)
names(pooled)[1]=c("inc")
pooled$inc<-pooled$inc/1000 #adjust scale
normal$case <- "normal case"
pooled$case <- "pooled, normal and special cases"
percentiles <- rbind(normal,pooled)

@@

<<specialcases11, fig.cap='Comparison of normal and special cases 2011',fig.pos='H'>>=

@@

# Export data to use with Stata
ns2011 <- rbind(data.frame(cbind("income" = pooled_i,"group"=0)),data.frame("income" = normal_i,"group"=1))
ns2011$group <- factor(ns2011$group,labels=c("pooled","normal"))
write.dta(ns2011, file="../../data/stata_data/ns2011.dta")

#Gini Range
#gini_allcases <- gini(ns2011$income[ns2011$group=="pooled"]); gini_allcases
#gini_normal <- gini(ns2011$income[ns2011$group=="normal"]); gini_normal
#round(gini_allcases-gini_normal,3)
# 0.02
# Es ist quatsch hier die interpolierten Reihen zu nehmen
# Gini 2011 Pooled (BH-Daten): 0.408
# Gini 2011 Normal (unsere ESTV Daten): 0.366
@

@@

# Export data to use with Stata
#habe_bd_2010_married <- rbind(data.frame(cbind("income" = habe10.married$inc_habe,"group"=0,"weight"=habe10.married$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
#habe_bd_2010_married$group <- factor(habe_bd_2010_married$group,labels=c("habe","bd"))
#write.dta(habe_bd_2010_married, file="../../data/stata_data/habe_bd_2010_married.dta")

# 2011
habe11.married<-habe0911 %>% 
  filter(Familientyp08>130,Jahr08==2011) %>% 
  mutate(psteink=Bruttoeinkommen08+A31m+A35m) 
# Wir brauchen (a) verfügbares Einkommen (b) Bruttoeinkommen und c(Bruttoeinkommen nach Soziversicherungsbeiträgen(A31m) und Monetäre Trasnferausgaben an andere HaushalteA35m)
habe11.married<-data.frame(habe11.married$VerfuegbaresEinkommen08*12/1000,habe11.married$Bruttoeinkommen08*12/1000,habe11.married$psteink*12/1000,habe11.married$jahresgewicht)
names(habe11.married)<-c("inc_habe","brutto","psteink","gew")
habe11.married$inc_habe[habe11.married$inc_habe<0]<-0 # One values is -500

bd.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(grepl("Verheiratet",Einheit))
start <- which(names(bd.2011)=="p1")
end <- which(names(bd.2011)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2011<- data.frame(t(bd.2011[1,start:end]),percentile)
names(bd.2011)[1]=c("inc")
bd.2011 <- rbind(c(0,0),bd.2011)
ival<-bd.2011$inc
perval<-bd.2011$percentile
bd.2011.ind<-unlist(sapply(2:length(ival),function(i){
  len=(perval[i]-perval[i-1])*100
  seq(ival[i-1],ival[i],length.out=len)
}))
bd.2011.ind<-bd.2011.ind/1000

habe_bd_2011_married <- rbind(data.frame(cbind("income" = habe11.married$inc_habe,"group"=0,"weight"=habe11.married$gew)),data.frame("income" = bd.2011.ind,"group"=1,"weight"=1))
habe_bd_2011_married$group <- factor(habe_bd_2011_married$group,labels=c("habe","bd"))
write.dta(habe_bd_2011_married, file="../../data/stata_data/habe_bd_2011_married.dta")

@@

estv_masse_long$measure <- factor(estv_masse_long$variable)
estv_masse_long$Year <- as.double(as.character(estv_masse_long$Year))
estv_masse_long <- estv_masse_long %>% group_by(measure) %>% mutate(log_value=log(value)-log(value)[1])
fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log_value,shape=measure))+
  #scale_y_continuous(limits=c(0,0.55))+
  scale_x_continuous(limits=c(1941,2011),breaks=number_ticks(10)) +
  geom_point(aes(shape=measure),size=3)+
  geom_line(aes(linetype=measure))+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Using different inequality measures")
fig_measures
